---
title: "Sink normalization"
author: "Beni Stocker"
date: "1/28/2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
df <- read_csv("data/T_Sland.csv", col_types = "ddddddddddddddd") %>% 
    pivot_longer(1:15, names_to = "source", values_to = "sland") %>% 
    left_join(
        read_csv("data/T_betaGPP.csv") %>% 
            pivot_longer(1:14, names_to = "source", values_to = "beta_gpp"),
        by = "source"
    ) %>% 
    left_join(
        read_csv("data/T_betaReco.csv") %>% 
            pivot_longer(1:13, names_to = "source", values_to = "beta_reco"),
        by = "source"
    ) %>% 
    left_join(
        read_csv("data/T_nonRespFlux.csv", col_types = "dddddddddddddd") %>% 
            pivot_longer(1:14, names_to = "source", values_to = "nrf"),
        by = "source"
    )
```

When reading the files you sent me, some data is missing.
```{r echo=FALSE}
knitr::kable(df)
```

Olot original Beta-GPP vs the sink. Note that this displays the 95% confidence interval to be somewhat larger than in ED Fig. 1.
```{r , warning=FALSE, message=FALSE}
df %>% 
    ggplot(aes(sland, beta_gpp)) +
    geom_point() +
    geom_smooth(method = "lm", color = "red", level = 0.95) +
    theme_classic()
```

Fit a linear regression model. Note that the R-squared is 0.95, not 0.99 as in the paper.
```{r}
linmod <- lm(sland ~ beta_gpp + beta_reco * nrf,
             data = df %>% 
                 filter(source != "GCPresidualSink")
             )
summary(linmod)
```

Apply variance normalization on the sink 
```{r}
cf <- coef(linmod)

df <- df %>% 
    mutate(betareco_nrf = beta_reco * nrf) %>% 
    mutate(b_beta_reco = cf["beta_reco"] * beta_reco,
           c_betareco_nrf = cf["beta_reco:nrf"] * betareco_nrf )

b_beta_reco_mean <- mean(df$b_beta_reco, na.rm = TRUE)
c_betareco_nrf_mean <- mean(df$c_betareco_nrf, na.rm = TRUE)

df <- df %>% 
    mutate(sland_norm = sland - ((b_beta_reco + c_betareco_nrf) - (b_beta_reco_mean + c_betareco_nrf_mean)))
```

Plot normalised sink vs beta-gpp. Doesn't look the way it should...
```{r, warning=FALSE, message=FALSE}
df %>% 
    ggplot(aes(sland_norm, beta_gpp)) +
    geom_point() +
    geom_smooth(method = "lm", color = "red", level = 0.95) +
    theme_classic()
```

Should it also be normalised for the non-respired flux (nrf)?
```{r}
cf <- coef(linmod)

df <- df %>% 
    mutate(d_nrf = cf["nrf"] * nrf )

d_nrf_mean <- mean(df$d_nrf, na.rm = TRUE)

df <- df %>% 
    mutate(sland_norm2 = sland - ((b_beta_reco + c_betareco_nrf + d_nrf) - (b_beta_reco_mean + c_betareco_nrf_mean + d_nrf_mean)))
```

Yes!
```{r, warning=FALSE, message=FALSE}
df %>% 
    ggplot(aes(sland_norm2, beta_gpp)) +
    geom_point() +
    geom_smooth(method = "lm", color = "red", level = 0.95) +
    theme_classic()
```

That's the one that gets the R-squared of 0.99:
```{r}
yardstick::rsq(df, sland_norm2, beta_gpp) %>% pull(.estimate)
```

Are the means of the original and normalised (red) sinks different? No!!!
```{r}
df %>% 
  ggplot() +
  geom_density(aes(sland_norm, ..density..)) +
  geom_density(aes(sland_norm2, ..density..), color = "red") +
  geom_vline(xintercept = mean(df$sland_norm, na.rm = TRUE), size = 2) +
  geom_vline(xintercept = mean(df$sland_norm2, na.rm = TRUE), color = "red") +
  theme_classic()
```


Now, perturb the distribution by removing a model (the first one: CABLE).
Let's make this a function.
```{r}
normalize_sink <- function(df){
  linmod <- lm(sland ~ beta_gpp + beta_reco * nrf,
             data = df %>% 
                 filter(source != "GCPresidualSink") %>% 
                 slice(-1)
             )
  cf <- coef(linmod)
  out <- df %>% 
      mutate(betareco_nrf = beta_reco * nrf) %>% 
      mutate(b_beta_reco = cf["beta_reco"] * beta_reco,
             c_betareco_nrf = cf["beta_reco:nrf"] * betareco_nrf,
             d_nrf = cf["nrf"] * nrf)
  
  b_beta_reco_mean <- mean(out$b_beta_reco, na.rm = TRUE)
  c_betareco_nrf_mean <- mean(out$c_betareco_nrf, na.rm = TRUE)
  d_nrf_mean <- mean(out$d_nrf, na.rm = TRUE)

  out <- out %>% 
    mutate(sland_norm2 = sland - ((b_beta_reco + c_betareco_nrf + d_nrf) - (b_beta_reco_mean + c_betareco_nrf_mean + d_nrf_mean)))
  
  return(out)
}

df2 <- df %>% 
  slice(-1) %>% 
  normalize_sink()
```


Plot both distributions and means: it does affect the mean (vertical lines).
```{r}
ggplot() +
  geom_density(aes(sland_norm2, ..density..), data = df) +
  geom_density(aes(sland_norm2, ..density..), data = df2, color = "red") +
  geom_vline(xintercept = mean(df$sland_norm2, na.rm = TRUE), lintetype = "dashed") +
  geom_vline(xintercept = mean(df2$sland_norm2, na.rm = TRUE), color = "red", lintetype = "dashed") +
  theme_classic()
```

